---
title: "NHANES Challenge"
format: html
editor: visual
---

### Load Libraries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
library(tidyverse)
library(haven)
library(here)
library(ggplot2)
library(janitor)
```

### Read-in Datasets for this Analysis:

-   DR1TOT_J.xpt as 'diet_behavior'
-   2017-2018_Hg-Blood.XPT as 'blood_hg'
-   2017-2018_Hg-Urine.XPT as 'urine_hg'
-   2017-2018_Diabetes.XPT as 'diabetes'

```{r}
diet_behavior <- read_xpt(here("data/nhanes_data/DR1TOT_J.XPT"))
blood_hg <- read_xpt(here("data/nhanes_data/2017-2018_Hg-Blood.XPT"))
urine_hg <- read_xpt(here("data/nhanes_data/2017-2018_Hg-Urine.XPT"))
diabetes <- read_xpt(here("data/nhanes_data/2017-2018_Diabetes.XPT"))
demographics <- read_xpt(here("data/nhanes_data/2017-2018_Demographics.XPT"))
```

### Subset Read-in Datasets

Subset 'diet_behavior' as 'diet'

```{r}
diet <- select(diet_behavior, SEQN, DRD360, DRD370B, DRD370BQ, DRD370Q, DRD370QQ)
```

Subset 'diabetes' as 'tiid'

```{r}
tiid <- select(diabetes, SEQN, DIQ010)
```

Subset 'blood_hg' as 'bhg'

```{r}
bhg <- select(blood_hg, SEQN, LBXIHG, LBDIHGSI, LBXBGE, LBXBGM)
```

Subset "urine_hg' as 'uhg'

```{r}
uhg <- select(urine_hg, SEQN, URXUHG)
```

### Merge Subsets Into A Working Dataframe as 'df'

```{r}
df <- list(diet, tiid, bhg, uhg)

df <- df %>% reduce(full_join, by = 'SEQN')
```

1.  Filter Dataframe df for the following:

```{r}
# Assuming your dataframe is named `nhanes_data`
df <- df %>%
  # Filter out rows where DIQ010 or DRD360 are NA
  filter(!is.na(DIQ010), !is.na(DRD370B)) %>%
  # Keep only rows where DIQ010 and DRD360 are 1 or 2
  filter(DIQ010 %in% c(1, 2), DRD370B %in% c(1, 2)) %>%
  # Recode 1 to "Yes" and 2 to "No" for DIQ010 and DRD360
  mutate(
    DIQ010 = ifelse(DIQ010 == 1, "Has Diabetes", "No Diabetes"),
    DRD370B = ifelse(DRD370B == 1, "Consumes Ahi", "No Ahi!")
  )
  
```

2.  Do some exploratory data analysis

```{r}
ggplot(data = df, aes(x = DRD360)) + 
  geom_bar() + 
  facet_grid(~DRD370B) +
  labs(title = "Fish Eaten in the Past 30 Days", x = "Fish Eaten" , y = "Count")
```

## Bayes Theorem - P(B\|A)

B \<- Has Diabetes

A \<- Consumes Ahi

2.  Determine the prior probability that someone has diabetes (DIQ010)

    -   P(B)

```{r}
# Prior Probabillity person has diabetes
df %>% 
  tabyl(DIQ010) 
```

```{r}
prior_probabillity_someone_has_diabetes = 0.1287574
```

2.  Determine the prior probability that someone eats tuna (DRD370B)
    -   P(A)

```{r}
# Prior Probabillity person eats tuna
df %>% 
  tabyl(DRD370B) # gets the number of types in a particular column


prior_probabillity_someone_eats_tuna = 0.3609467
```

2.  Determine the likelihood of eating tuna and having diabetes
    -   L(B\|A) = P(B\|A)

```{r}
 df %>% 
   tabyl(DRD370B,DIQ010 ) %>% 
   adorn_percentages("col")
 
liklihood_someone_eats_tuna_has_diabetes = 0.3584559
```

2.  Determine posterior probability via bayes theorem
    -   P(B\|A)

```{r}
(prior_probabillity_someone_has_diabetes * liklihood_someone_eats_tuna_has_diabetes) / prior_probabillity_someone_eats_tuna
```

2.  What other questions can we ask of this data?

We can ask if someone has diabetes if they eat swordfish. We can also do if they don't eat tuna or swordfish.

#### Shorter Method for Bayes Theorem

P(A\|B)

```{r}
df %>% 
  tabyl(DIQ010, DRD370B) %>% 
  adorn_percentages("col")
```

.13 would be our posterior probabillity

#### Diabetes & Age

Selecting Columns from demographics column

```{r}
demographics <- demographics %>% 
  select(SEQN, RIDRETH1, RIAGENDR, RIDAGEYR)
```

Joining 2 data frames (df and demographics) together by "SEQN" Column

```{r}
nhanes_data <- left_join(df, demographics, by = "SEQN")
```

Creating a new column "Age Bracket"

```{r}
nhanes_data <- nhanes_data %>% 
  mutate(age_bracket = ifelse(RIDAGEYR > 40,"Over 40", "Under 40"))
```

#### Density Plot

```{r}
ggplot(nhanes_data, aes(x = RIDAGEYR, fill = DIQ010)) +
  geom_density(alpha = 0.5) +
  labs(title = "Has Diabetes based on Age", x = "Age", y = "Density")
```

P(B\|A) Finding the probability that someone has diabetes given their age is over 40

B \<- has diabetes

A \<- Age over 40

P(B)

```{r}
nhanes_data %>% 
  tabyl(DIQ010)
```

```{r}
prior_probabillity_has_diabetes_nhanes <- 0.1287574
```

-   L(B\|A) = P(B\|A)

```{r}
nhanes_data %>%
  tabyl(age_bracket, DIQ010) %>% 
  adorn_percentages("col")
  
```

```{r}
likelhood_someone_has_diabetes_over_40 <- 0.95588235
```

P(A)

```{r}
nhanes_data %>%
  tabyl(age_bracket)
```

```{r}
prior_probability_someone_is_over_40 <- 0.5342012
```

Finding the posterior probability

```{r}
(prior_probabillity_has_diabetes_nhanes * likelhood_someone_has_diabetes_over_40) / prior_probability_someone_is_over_40
```

Shorter Methods for Bayes Theorem P(B\|A)

```{r}
nhanes_data %>% 
  tabyl(DIQ010, age_bracket) %>% 
  adorn_percentages("col")
```
